# Simulation

Here we mainly have 4 steps.
-    Data Generation.
-    Initialization.
-    Gibbs.
-    Prediction.

As we have finished step 1 and 3 before, here only initialization and prediction.

## Initialization
```{r}
#' Unified Initialization for Latent Variables
#' 
#' @param type String. Must be either "ordinal" or "nominal".
#' @param params A list containing the required parameters for initialization.
#' @param seed Optional random seed to ensure reproducibility.
#' 
#' @return A list containing initial values (e.g., U0, Tau0).
gibbs.init <- function(type, params, seed = NULL) {
  
  # Validate type input
  if (!(type %in% c("ordinal", "nominal"))) {
    stop("Input Error: 'type' must be 'ordinal' or 'nominal'.")
  }
  
  # Set seed if provided
  if (!is.null(seed)) {
    set.seed(seed)
    params$seed <- NULL 
  }
  
  # Dispatch to the specific initialization function
  if (type == "ordinal") {
    return(do.call(gibbs.init.ordinal, params))
  } else {
    return(do.call(gibbs.init.nominal, params))
  }
}

#' Generate Initial Values for Ordinal Latent Variables
#' 
#' @param n Number of samples.
#' @param m Number of categories.
#' @param c Vector of category labels (1, 2, ..., m).
#' @param u.obs.idx Indices of the observed latent variables.
#' @param U.obs Vector of observed latent variable values.
#' @param u.lb Lower bound for latent variables (default -Inf).
#' @param u.ub Upper bound for latent variables (default Inf).
#' 
#' @return A list containing U0 (initial U), Tau0 (initial thresholds), and Tau.aug0.
gibbs.init.ordinal <- function (n, m, c, u.obs.idx, U.obs, u.lb = -Inf, u.ub = Inf) {
  
  # Initialize latent variable vector
  U <- rep(0, n)
  
  if (length(u.obs.idx) > 0) {
    # Case: Some latent variables are observed
    Tau.aug <- c(u.lb, rep(0, m - 1), u.ub)
    
    # Randomly draw initial thresholds consistent with observed data bounds
    for (j in 1:(m - 1)) {
      lower <- max(Tau.aug[j], U.obs[c[u.obs.idx] <= j])
      upper <- min(u.ub, U.obs[c[u.obs.idx] > j])
      Tau.aug[j + 1] <- DrawTrunNormal(0, 1, lower, upper)
    }
    
    # Generate initial U based on categories and thresholds
    U <- DrawTrunNormal(0, 1, Tau.aug[c], Tau.aug[c + 1])
    U[u.obs.idx] <- U.obs  # Fix the observed parts
    Tau <- Tau.aug[2:m]
    
  } else {
    # Case: No latent variables are observed
    # Generate sorted random thresholds
    Tau <- sort(rnorm(m - 1, u.lb, u.ub))
    Tau.aug <- c(u.lb, Tau, u.ub)
    U <- DrawTrunNormal(0, 1, Tau.aug[c], Tau.aug[c + 1])
  }
  
  return(list(U0 = U, Tau0 = Tau, Tau.aug0 = Tau.aug))
}

#' Generate Initial Values for Nominal Latent Variables
#' 
#' @param n Number of samples.
#' @param m Number of categories.
#' @param c Vector of category labels (1, 2, ..., m).
#' @param u.obs.idx Indices of the observed latent variable rows.
#' @param U.obs Matrix of observed latent variable values.
#' 
#' @return A list containing U0 (initial matrix of latent variables).
gibbs.init.nominal <- function (n, m, c, u.obs.idx, U.obs) {
  
  # Get dimensionality of latent space
  d <- ncol(U.obs)
  U <- matrix(0, n, d)
  
  # Fit a category-specific normal model based on observed U
  for (j in 1:m) {
    # Extract observed u belonging to category j
    obs_in_j_idx <- (c[u.obs.idx] == j)
    U.obs.j <- U.obs[obs_in_j_idx, , drop = FALSE]
    
    if (nrow(U.obs.j) > 1) {
      # Compute empirical mean and covariance matrix for category j
      mu.j <- apply(U.obs.j, 2, mean)
      U.centered <- t(t(U.obs.j) - mu.j)
      Sigma.j <- (t(U.centered) %*% U.centered) / sum(obs_in_j_idx)
      
      # Fill initial U for all samples in category j using this distribution
      U[c == j, ] <- rmvnorm(sum(c == j), mu.j, Sigma.j)
    } else {
      # Fallback if no/one observation exists for category j
      U[c == j, ] <- matrix(rnorm(sum(c == j) * d), ncol = d)
    }
  }
  
  # Ensure observed indices keep their true values
  U[u.obs.idx, ] <- U.obs
  
  return(list(U0 = U))
}

```


## Prediction
```{r}
#' Unified Oracle Predictor (Using True Parameters)
#' 
#' @param type String. Must be either "ordinal" or "nominal".
#' @param params A list containing the true parameters and test data.
#' @param seed Optional random seed for reproducibility.
#' 
#' @return A list containing histories of U*, y.predict, and f.predict.
gibbs.oracle <- function(type, params, seed = NULL) {
  
  # Validate type input
  if (!(type %in% c("ordinal", "nominal"))) {
    stop("Input Error: 'type' must be 'ordinal' or 'nominal'.")
  }
  
  # Set seed if provided
  if (!is.null(seed)) {
    set.seed(seed)
    params$seed <- NULL 
  }
  
  # Dispatch to the specific oracle function
  if (type == "ordinal") {
    return(do.call(gibbs.oracle.ordinal, params))
  } else {
    return(do.call(gibbs.oracle.nominal, params))
  }
}

#' Oracle Predictor for Ordinal Models
#' 
#' @param X.star Matrix of new quantitative predictors (test points).
#' @param c.star Vector of new ordinal categories.
#' @param beta True coefficient vector for u ~ X.
#' @param sigma.u2 True scalar variance for latent u.
#' @param sigma.error2 True variance for response error.
#' @param Tau True threshold vector.
#' @param u.lb Lower bound for latent u.
#' @param u.ub Upper bound for latent u.
#' @param ResponseFunction Function to calculate f(x, u).
#' @param code Design code passed to ResponseFunction.
#' @param uniform Logical. If TRUE, design of u is uniform.
#' @param n.mc Number of Monte Carlo samples.
#' 
#' @return A list containing U.star.hist, y.predict.hist, and f.predict.hist.
gibbs.oracle.ordinal <- function (X.star, c.star, beta, sigma.u2, sigma.error2, 
                                  Tau, u.lb, u.ub, ResponseFunction, code,
                                  uniform = FALSE, n.mc = 1000) {
  # Initialize
  n.star <- length(c.star)
  U.star.hist <- matrix(0, n.star, n.mc)
  f.predict.hist <- matrix(0, n.star, n.mc)
  y.predict.hist <- matrix(0, n.star, n.mc)
  X.star.tilde <- cbind(1, X.star)
  
  for (idx in 1:n.mc) {
    # 1. Draw U.star
    Tau.aug <- c(u.lb, Tau, u.ub)
    if (uniform) {
      U.star <- runif(n.star, Tau.aug[c.star], Tau.aug[c.star+1])
    } else {
      U.star <- DrawTrunNormal(X.star.tilde %*% beta, rep(sqrt(sigma.u2), n.star), 
                               Tau.aug[c.star], Tau.aug[c.star+1])
    }
    U.star.hist[, idx] <- U.star
    
    # 2. Draw f.star and y.star
    f.predict.hist[, idx] <- ResponseFunction(X.star, as.matrix(U.star), code)
    y.predict.hist[, idx] <- f.predict.hist[, idx] + rnorm(n.star, 0, sqrt(sigma.error2))
  }
  
  return(list(U.star.hist = U.star.hist,
              y.predict.hist = y.predict.hist, 
              f.predict.hist = f.predict.hist))
}

#' Oracle Predictor for Nominal Models
#' 
#' @param X.star Matrix of new quantitative predictors (test points).
#' @param c.star Vector of new nominal categories.
#' @param beta True coefficient matrix (B.tilde) for u ~ X.
#' @param sigma.u2 True covariance matrix (Sigma.u) for latent u.
#' @param sigma.error2 True variance for response error.
#' @param classifier A function returning the category c given u (the Error Mechanism).
#' @param ResponseFunction Function to calculate f(x, u).
#' @param code Design code passed to ResponseFunction.
#' @param uniform Logical. If TRUE, design of u is uniform.
#' @param u.lb,u.ub Bounds for uniform design.
#' @param n.mc Number of Monte Carlo samples.
#' 
#' @return A list containing U.star.hist, y.predict.hist, and f.predict.hist.
gibbs.oracle.nominal <- function (X.star, c.star, beta, sigma.u2, sigma.error2, 
                                  classifier, ResponseFunction, code,
                                  uniform = FALSE, u.lb = NULL, u.ub = NULL, n.mc = 1000) {
  # Initialize
  n.star <- length(c.star)
  d <- ncol(sigma.u2)
  U.star.hist <- matrix(0, n.star, d * n.mc)
  f.predict.hist <- matrix(0, n.star, n.mc)
  y.predict.hist <- matrix(0, n.star, n.mc)
  
  X.star.tilde <- cbind(1, X.star)
  mean.u <- X.star.tilde %*% beta
  Chol.u <- t(chol(sigma.u2))
  
  for (idx in 1:n.mc) {
    U_current_mc <- matrix(0, n.star, d)
    
    # 1. Draw U* using Accept-Reject based on the classifier
    for (i in 1:n.star) {
      while (TRUE) {
        if (!uniform) {
          U.prime <- c(mean.u[i, ] + Chol.u %*% rnorm(d))
        } else {
          U.prime <- runif(d, u.lb, u.ub)
        }
        
        # Check if the drawn U belongs to the observed category c.star
        c.prime <- classifier(U.prime)$c
        if (c.prime == c.star[i]) {
          U_current_mc[i, ] <- U.prime
          break
        }
      }
    }
    U.star.hist[, ((idx-1)*d+1):(idx*d)] <- U_current_mc
    
    # 2. Draw f.star and y.star based on accepted U*
    f.predict.hist[, idx] <- ResponseFunction(X.star, U_current_mc, code)
    y.predict.hist[, idx] <- f.predict.hist[, idx] + rnorm(n.star, 0, sqrt(sigma.error2))
  }
  
  return(list(U.star.hist = U.star.hist,
              y.predict.hist = y.predict.hist, 
              f.predict.hist = f.predict.hist))
}

```

## Ordinal Example

### Data generation
Set basic parameters for simulation.

```{r}
n <- 100  # sample size
m <- 6  # number of classes
d.x <- 1  # dimension of x

# true parameters for the data-generating process
sigma.error2.true <- 0.1^2
beta.true <- c(pi, 0)
sigma.u2.true <- (0.5*pi)^2
```

Generate other parameters and data.

```{r}
set.seed(1)
code <- 'cos'  # code for the response function
# generate x, u, c ----
x.ub <- 2
x.lb <- -2
X <- (x.ub - x.lb)*maximinLHS(n, d.x) + x.lb 
X.tilde <- cbind(1, X)
u.lb <- 0
u.ub <- 2*pi
U <- matrix(DrawTrunNormal(X.tilde%*%beta.true, sqrt(sigma.u2.true), u.lb, u.ub), n, 1)
#U <- matrix(X.tilde%*%beta.true+sqrt(sigma.u2.true)*rnorn(n), n, 1)
Tau.true <- (1:(m-1))/m*(u.ub-u.lb) + u.lb
c <- apply(U, 1, function(u) match(T, c(u-Tau.true,-Inf) <= 0))  
# generate y ----
f.true <- ResponseFunction(X, U, code)
y <- f.true + sqrt(sigma.error2.true)*rnorm(n) 
sd.y <- sd(y)
y <- y/sd.y  # standardize the response
# select observed u ----
n.obs <- floor(10)  # total number
n.obs.c <- rep(0, m)  # number for each class
n.min <- 2
for (i in 1:m) {
  if (i < m) {
    n.obs.c[i] <- max(floor(sum(c == i)/n*n.obs), n.min)
  } else {
    n.obs.c[i] <- max(n.obs - sum(n.obs.c[-m]), n.min)
  }
}
u.obs.idx <- NULL
for (i in 1:m) {
  u.obs.idx <- c(u.obs.idx, sample(c(1:n)[c == i], n.obs.c[i], FALSE))
}
u.mis.idx <- c(1:n)[-u.obs.idx]
U.obs <- U[u.obs.idx, ]
n.obs <- sum(n.obs.c)
```


### Visualization

```{r fig.cap = 'True response function', echo=FALSE}
PlotResponse(code = code, nx = 100, lb.u = min(U), ub.u = max(U), lb.x = min(X), ub.x = max(X))
```

```{r}
u_df <- data.frame(x = X, u = U, y = y*sd.y, class = as.character(c), observed = rep("FALSE", n))
u_df$observed[u.obs.idx] <- "TRUE"
```

```{r fig.cap = 'Scatterplot of y vs. u', echo=FALSE}
ggfig <- 
ggplot(u_df, aes(x = u, color = class)) + 
  geom_point(size = 3.5, aes(y = y)) + 
  scale_color_viridis_d(name = "Class") + 
  geom_function(fun = function(u) ResponseFunction(X=0, U=u, code=code), color = '#FF7F0E', linewidth = 1) +
  geom_vline(xintercept = Tau.true, linewidth = 1) + 
  coord_cartesian(xlim = c(min(U), max(U)), ylim = c(min(u_df$y), max(u_df$y))) +
  labs(color = "Class", x = 'Latent u') + 
  my_theme
#ggsave(paste0(params$fig.dir, 'Scatter-y-vs-u.pdf'), plot = ggfig, width = 8, height = 6)

ggfig <- 
ggplot(u_df[u_df$observed=="TRUE",], aes(x = u, y = y, color = class)) +
  geom_point(size = 3.5) +
  scale_color_viridis_d(name = "Class") + 
  geom_function(fun = function(u) ResponseFunction(X=0, U=u, code=code), color = '#FF7F0E', linewidth = 1) +
  coord_cartesian(xlim = c(min(U), max(U)), ylim = c(min(u_df$y), max(u_df$y))) +
  labs(color = "Class") + 
  my_theme
#ggsave(paste0(params$fig.dir, 'Scatter-y-vs-u-obs.pdf'), plot = ggfig, width = 8, height = 6)

ggfig <- 
ggplot(u_df, aes(x = c, y = y, color = class)) +
  geom_point(size = 3.5) +
  scale_color_viridis_d(name = "Class") + 
  labs(color = "Class", x = 'Class') + 
  my_theme
#ggsave(paste0(params$fig.dir, 'Scatter-y-vs-c.pdf'), plot = ggfig, width = 8, height = 6)
```

```{r fig.cap = 'Scatterplot of u versus x', echo=FALSE}
ggfig <- 
  ggplot(u_df, aes(x = X, y = u, color = class, shape = observed)) +
  geom_point(size = 3) +
  scale_color_viridis_d(name = "Class") + 
  geom_abline(slope = beta.true[2], intercept = beta.true[1], linewidth = 1, color = 'brown') +
  labs(color = "Class", shape = 'Observation') + 
  my_theme
#ggsave(paste0(params$fig.dir, 'Scatter-u-vs-x.pdf'), plot = ggfig, width = 8, height = 6)
```

```{r fig.cap = 'Barplot of c', echo=FALSE}
c_df <- data.frame(class = 1:m, value = rep(0,m))
for (i in 1:m) {
  c_df$value[i] <- sum(c == i)
}
bar_plot <- 
  ggplot(c_df, aes(x = class, y = value, fill = as.character(class))) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  scale_fill_viridis_d(name = "Class") +  
  labs(x = "Class", y = "Count") +
  my_theme 
#ggsave(paste0(params$fig.dir, 'class-bar-plot.pdf'), plot = bar_plot, width = 8, height = 6)
```

```{r fig.cap = 'Scatterplot of y vs. x-c-u-obs', echo=FALSE}
ggfig <- Plot.Data(c, X, y, U, u.obs.idx)
ggsave(paste0(params$fig.dir, 'Scatter-y-x-c.pdf'), plot = ggfig[[1]], width = 8, height = 6)
ggsave(paste0(params$fig.dir, 'Scatter-y-x-u-obs.pdf'), plot = ggfig[[2]], width = 8, height = 6)
```

### Hyperparameters Related

fixed hyperparameter

```{r, eval = FALSE}
v0 <- 1
g <- 1
IG.a.u <- 1
IG.b.u <- 1
IG.a.error <- 2.5/2
IG.b.error <- 2.5*0.084/2
```

Obtain GP hyperparameters by MLE

Fit MLE for the GP hyperparameters using the limited observed data. We use package 'laGP' (see Gramacy's book).

```{r}
W <- cbind(X, U)
gp <- newGPsep(W[u.obs.idx,], y[u.obs.idx], d = 0.1, g = 0.01, dK = TRUE)
gp.mle <- mleGPsep(gp, param = "both", tmin = c(1e-3, 1e-3), tmax=c(1e3, 10^2))
theta.mle.obs <- 1/gp.mle$theta[1:(d.x+1)]
g.mle <- gp.mle$theta[length(gp.mle$theta)] 
sigma2.mle <- (t(y[u.obs.idx])%*%solve(KernelMatrix(W[u.obs.idx,], c(theta.mle.obs, 1)) + g.mle*diag(n.obs))%*%y[u.obs.idx]/n.obs) |> drop()
sigma.error2.mle.obs <- sigma2.mle*g.mle
rho.mle.obs <- sqrt(1/g.mle)
```

```{r echo = FALSE}
print(paste0('The MLE (from observed U) of lengthscale for x is ', theta.mle.obs[1:d.x]))
print(paste0('The MLE (from observed U) of lengthscale for u is ', theta.mle.obs[d.x+1]))
print(paste0('The MLE (from observed U) of nugget (variance) is ', sigma.error2.mle.obs))
print(paste0('The MLE (from observed U) of SN ratio is ', rho.mle.obs))
```

```{r}
# call the destructor function otherwise memory will leak.
deleteGPsep(gp)
```

#### (Optional) For comparison, we can also obtain the MLE of GP hyperparameters using the complete data. 

```{r, eval = FALSE}
W <- cbind(X, U)
gp <- newGPsep(W, y, d = 0.1, g = 0.01, dK = TRUE)
gp.mle <- mleGPsep(gp, param = "both", tmin = c(1e-3, 1e-3), tmax=c(1e3, var(y)))
theta.mle <- 1/gp.mle$theta[1:(d.x+1)]
g.mle <- gp.mle$theta[length(gp.mle$theta)] 
sigma2.mle <- (t(y)%*%solve(KernelMatrix(W, c(theta.mle, 1)) + g.mle*diag(n))%*%(y)/n) |> drop()
sigma.error2.mle <- sigma2.mle*g.mle
rho.mle <- sqrt(1/g.mle)
rm(list=c('W','gp','gp.mle','g.mle','sigma2.mle'))
```

```{r echo = FALSE, eval = FALSE}
print(paste0('The MLE of lengthscale for x is ', theta.mle[1:d.x]))
print(paste0('The MLE of lengthscale for u is ', theta.mle[d.x+1]))
print(paste0('The MLE of nugget (variance) is ', sigma.error2.mle))
print(paste0('The MLE of SN ratio is ', rho.mle))
```

#### (Optional) Obtain GP hyperparameters by Empirical Bayes

Alternatively, we can obtain GP hyperparameters by EB, which is much slower than the MLE approach using the observed $u$.

Create candidates for GP hyperparameters (by grid).

```{r, eval = FALSE}
hyper.mat <- c() # each column of hyper.mat is a candidate of hyperparameter (theta,rho)
for (i in exp(log(2)*c(-6:1))) { # length scale for x
  for (j in exp(log(2)*c(-6:1))) { # lengthscale for u
    for(l in exp(log(2)*c(2:5))) { # signal-to-noise rho
      hyper.mat <- cbind(hyper.mat, c(rep(i, d.x), j, l))
    }
  }
}
rm(list=c('i','j','l'))
```

Generate Monte Carlo samples of $u$ given $\{x,c,u_\text{obs}\}$.

```{r, eval = FALSE}
n.mc.eb <- 10000
sigma.u20 <- 1/rgamma(1, shape = IG.a.u, rate = IG.b.u)
beta0 <- rep(0, d.x+1)
Ini.par <- GetInitial(n, m, c, u.obs.idx, U.obs, u.lb, u.ub) 
Tau0 <- Ini.par$Tau0
U0 <- Ini.par$U0
samples.eb <- Gibbs.EB(seed = 345, n.mc = n.mc.eb, 
                       beta0 = beta0, sigma.u20 = sigma.u20, Tau0 = Tau0, U0 = U0, 
                       X = X, c = c, u.obs.idx = u.obs.idx, U.obs = U.obs, u.lb = u.lb, u.ub = u.ub,
                       v0 = v0, g = g, IG.a.u = IG.a.u, IG.b.u = IG.b.u)
```

Estimate the marginal likelihood (using parallel computing).

```{r, eval = FALSE}
cl <- parallel::makeCluster(8)
doParallel::registerDoParallel(cl)
eb.evidence <- foreach(i = 1:ncol(hyper.mat), .combine = 'cbind') %dopar% {
  EstimateEvidence(theta = hyper.mat[1:(d.x+1), i], rho = hyper.mat[d.x+1+1, i],
                   U.hist = samples.eb$U.hist,
                   IG.a.error = IG.a.error, IG.b.error = IG.b.error,
                   X = X, y = y)
}
parallel::stopCluster(cl)
rm(cl)
```

Select the optimal one.

```{r, eval = FALSE}
skip <- 10
eb.evidence.mean <- apply(eb.evidence[seq(1, nrow(eb.evidence), by = skip), ], 2, mean)
eb.evidence.sd <- apply(eb.evidence[seq(1, nrow(eb.evidence), by = skip), ], 2, sd)
theta.eb <- hyper.mat[1:(d.x+1), which.max(eb.evidence.mean)]
rho.eb <- hyper.mat[d.x+1+1, which.max(eb.evidence.mean)]
```

Plot the Monte Carlo estimates of the marginal likelihood for each candidate. 

```{r echo = FALSE, eval = FALSE}
plot(log(eb.evidence.mean), 
     ylab = 'Log Evidence', 
     xlab = 'Hyper-parameter index')
     #ylim = c(min(eb.evidence.mean - 2*eb.evidence.sd/sqrt(params$n.mc)), 
     #         max(eb.evidence.mean + 2*eb.evidence.sd/sqrt(params$n.mc))))
points(which.max(eb.evidence.mean), log(max(eb.evidence.mean)), col = 'blue', pch = 16)
#segments(1:ncol(hyper.mat), eb.evidence.mean - 2*eb.evidence.sd/sqrt(params$n.mc),
#         1:ncol(hyper.mat), eb.evidence.mean + 2*eb.evidence.sd/sqrt(params$n.mc))
```

```{r echo = FALSE, eval = FALSE}
plot(eb.evidence.mean, 
     ylab = 'Evidence', 
     xlab = 'Hyper-para index',
     ylim = c(min(eb.evidence.mean - 2*eb.evidence.sd/sqrt(n.mc.eb)), 
              max(eb.evidence.mean + 2*eb.evidence.sd/sqrt(n.mc.eb))), pch = 16, cex = 0.5)
segments(1:ncol(hyper.mat), eb.evidence.mean - 2*eb.evidence.sd/sqrt(n.mc.eb),
         1:ncol(hyper.mat), eb.evidence.mean + 2*eb.evidence.sd/sqrt(n.mc.eb))
points(which.max(eb.evidence.mean), max(eb.evidence.mean), col = 'blue', pch = 16)
```

### Bayesian Inference by Gibbs

Initialize.

```{r}
set.seed(999)
beta.hist <- matrix(rep(0, d.x+1), d.x+1, 1)
sigma.u2.hist <- 1/rgamma(1, shape = IG.a.u, rate = IG.b.u)
sigma.error2.hist <- 1/rgamma(1, shape = IG.a.error, rate = IG.b.error)
Ini.par <- gibbs.init.ordinal(n, m, c, u.obs.idx, U.obs, u.lb, u.ub) 
Tau.hist <- matrix(Ini.par$Tau0, m-1, 1)
U.hist <- matrix(Ini.par$U0, n, 1)
accep.hist <- NULL
f.hist <- NULL

K = 1000  # number of iteration for a grand cycle
lmd <- 0.01  # step size, adjusted after each cycle 
stepsize.hist <- data.frame(stepsize = lmd, k = 1)
time0 <- proc.time()
time.hist <- data.frame(user = double(), system = double(), elapsed = double())
```

Run the Gibbs sampler.

```{r}
for (cycles in 1:60) {
  # Initialize the current cycle
  beta0 <- beta.hist[, ncol(beta.hist)]
  sigma.u20 <- sigma.u2.hist[length(sigma.u2.hist)]
  sigma.error20 <- sigma.error2.hist[length(sigma.error2.hist)]
  Tau0 <- Tau.hist[, ncol(Tau.hist)]
  U0 <- U.hist[, ncol(U.hist)]
  
  # Run Gibbs for the current cycle 
  Gibbs1 <- gibbs.ordinal(num.iter = K, lmd = lmd, 
                  beta0 = beta0, sigma.u20 = sigma.u20, sigma.error20 = sigma.error20, Tau0 = Tau0, U0 = U0, 
                  X = X, c = c, y = y, u.obs.idx = u.obs.idx, U.obs = U.obs, u.lb = u.lb, u.ub = u.ub,
                  v0 = v0, g = g, IG.a.u = IG.a.u, IG.b.u = IG.b.u, IG.a.error = IG.a.error, IG.b.error = IG.b.error,
                  theta = theta.mle.obs, rho = rho.mle.obs, simu.f = FALSE, design.map = FALSE) # theta = theta.eb, rho = rho.eb
  
  # save results
  beta.hist <- cbind(beta.hist, Gibbs1$beta.hist)
  sigma.u2.hist <- c(sigma.u2.hist, Gibbs1$sigma.u2.hist)
  sigma.error2.hist <- c(sigma.error2.hist, Gibbs1$sigma.error2.hist)
  Tau.hist <- cbind(Tau.hist, Gibbs1$Tau.hist)
  U.hist <- cbind(U.hist, Gibbs1$U.hist)
  accep.hist <- c(accep.hist, Gibbs1$accep)
  f.hist <- cbind(f.hist, Gibbs1$f.hist)
  
  # adjust stepsize and record time
  if (mean(Gibbs1$accep) <= 0.2) {
    lmd <- lmd / 1.5
    stepsize.hist[nrow(stepsize.hist) + 1, ] <- c(lmd, K*cycles+1)
  }else if (mean(Gibbs1$accep) > 0.5){
    lmd <- lmd * 1.5
    stepsize.hist[nrow(stepsize.hist) + 1, ] <- c(lmd, K*cycles+1)
  }
  time.hist[cycles, ] <- c(proc.time() - time0)[1:3]
}
```

### Save results 

```{r}
saveRDS(list(X = X, y = y, sd.y = sd.y, c = c, x.ub = x.ub, x.lb = x.lb,
             v0 = v0, g = g, IG.a.u = IG.a.u, IG.b.u = IG.b.u, IG.a.error = IG.a.error, IG.b.error = IG.b.error, u.lb = u.lb, u.ub = u.ub,
             theta.mle.obs = theta.mle.obs, rho.mle.obs = rho.mle.obs,
             #theta.eb = theta.eb, rho.eb = rho.eb,
             beta.hist = beta.hist, beta.true = beta.true,
             sigma.u2.hist = sigma.u2.hist, sigma.u2.true = sigma.u2.true,
             sigma.error2.hist = sigma.error2.hist, sigma.error2.true = sigma.error2.true,
             Tau.hist = Tau.hist, Tau.true = Tau.true,
             U.hist = U.hist, U.true = U, u.obs.idx = u.obs.idx, U.obs = U.obs,
             f.hist = f.hist, f.true = f.true,
             accep.hist = accep.hist, stepsize.hist = stepsize.hist, time.hist = time.hist), 
        file = params$save.filename) 
```

Shutdown the computer.

```{r echo = FALSE}
# system('shutdown -s')
```